---
title: "Week 2 - Homework"
author: "Omar Boffil"
date: '05/30/2020'
output:
  pdf_document: default
  html_document: 
    theme: readable
    toc: yes  
urlcolor: cyan
---


***

## Exercise 1 (Using `lm`)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Suppose we would like to understand the size of a cat's heart based on the body weight of a cat. Fit a simple linear model in `R` that accomplishes this task. Store the results in a variable called `cat_model`. Output the result of calling `summary()` on `cat_model`.

```{r}
library(MASS)
body = cats$Bwt
heart = cats$Hwt
cat_model = lm(heart ~ body, data = cats)
summary(cat_model)
```

**(b)** Output only the estimated regression coefficients. Interpret $\hat{\beta_0}$ and $\beta_1$ in the *context of the problem*. Be aware that only one of those is an estimate.
```{r}
cat_model$coefficients
```
$\hat{\beta_1}$ of 4.0341 g is the estimated increase in mean for the cat's heart weight to the rise of 1 kg cat weight. 

$\hat{\beta_0}$ is -0.3566624 g is the estimate mean cat's heart weight for a cat of 0 kg weight

**(c)** Use your model to predict the heart weight of a cat that weights **3.1** kg. Do you feel confident in this prediction? Briefly explain.

```{r}
predict(cat_model, newdata = data.frame(body = 3.1))
```
Yes, I feel confident because it is close to the mean which is 12.5

**(d)** Use your model to predict the heart weight of a cat that weights **1.5** kg. Do you feel confident in this prediction? Briefly explain.

```{r}
predict(cat_model, newdata = data.frame(body = 1.5))
```
Yes, because the values is close to the mean for that value

**(e)** Create a scatterplot of the data and add the fitted regression line. Make sure your plot is well labeled and is somewhat visually appealing.
```{r}
plot(heart ~ body, data = cats, main = "Cat's heart weight based on the body weight",
     xlab = "Body weight", ylab = "Heart weight", col = "red", pch  = 20 
     )

abline(lm(heart ~ body, data = cats), col = "blue")
```

**(f)** Report the value of $R^2$ for the model. Do so directly. Do not simply copy and paste the value from the full output in the console after running `summary()` in part **(a)**.

$R^2$ is `r summary(cat_model)$r.squared`


***

## Exercise 2 (Writing Functions)

This exercise is a continuation of Exercise 1.

**(a)** Write a function called `get_sd_est` that calculates an estimate of $\sigma$ in one of two ways depending on input to the function. The function should take three arguments as input:

- `fitted_vals` - A vector of fitted values from a model
- `actual_vals` - A vector of the true values of the response
- `mle` - A logical (`TRUE` / `FALSE`) variable which defaults to `FALSE`

The function should return a single value:

- $s_e$ if `mle` is set to `FALSE`.
- $\hat{\sigma}$ if `mle` is set to `TRUE`.

```{r}
get_sd_est = function(fitted_vals, actual_vals, mle = FALSE){
  
  if (mle == FALSE){
    summary(fitted_vals)$sigma
  }else{
    sd(actual_vals - mean(actual_vals)) 
  }
  
}
```

**(b)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `FALSE`. Explain the resulting estimate in the context of the model.
```{r}
get_sd_est(cat_model, cats$Hwt, FALSE)
```


**(c)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `TRUE`. Explain the resulting estimate in the context of the model. Note that we are trying to estimate the same parameter as in part **(b)**.
```{r}
get_sd_est(cat_model, cats$Hwt, TRUE)
```

**(d)** To check your work, output `summary(cat_model)$sigma`. It should match at least one of **(b)** or **(c)**.
```{r}
all.equal(get_sd_est(cat_model, cats$Hwt, FALSE), summary(cat_model)$sigma)
all.equal(get_sd_est(cat_model, cats$Hwt, TRUE), summary(cat_model)$sigma)
```

***

## Exercise 3 (Simulating SLR)

Consider the model

\[
Y_i = 5 + -3 x_i + \epsilon_i
\]

with 

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 10.24)
\]

where $\beta_0 = 5$ and $\beta_1 = -3$.

This exercise relies heavily on generating random observations. To make this reproducible we will set a seed for the randomization. Alter the following code to make `birthday` store your birthday in the format: `yyyymmdd`. For example, [William Gosset](https://en.wikipedia.org/wiki/William_Sealy_Gosset), better known as *Student*, was born on June 13, 1876, so he would use:

```{r}
birthday = 19900826
set.seed(birthday)
```

**(a)** Use `R` to simulate `n = 25` observations from the above model. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
num_obs = 25
beta_0 = 5
beta_1 = -3
sigma = sqrt(10.24)
set.seed(2)
above = rnorm(25, 0, sigma)
above
set.seed(1)
x = runif(n = 25, 0, 10)
```

You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Store the data frame this function returns in a variable of your choice. Note that this function calls $y$ `response` and $x$ `predictor`.

**(b)** Fit a model to your simulated data. Report the estimated coefficients. Are they close to what you would expect? Briefly explain.

```{r}
y = beta_0 + beta_1 * x + above
df_fit = lm(y ~ x)
coef(df_fit)
```
The coefficients are what I expect they are close in value to the actual parameters

**(c)** Plot the data you simulated in part **(a)**. Add the regression line from part **(b)** as well as the line for the true model. Hint: Keep all plotting commands in the same chunk.

```{r}
plot(y ~ x, main = "Simulated Regression Data",
     xlab = "Predictor",
     ylab = "Response")
abline(df_fit, col = "blue")
abline(beta_0, beta_1, lty = 2, col = "red")
legend("topright", c("Estimate", "Truth"), lty = c(1, 2), lwd = 2, col = c("blue", "red"))
```


**(d)** Use `R` to repeat the process of simulating `n = 25` observations from the above model $1500$ times. Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. Some hints:

- Consider a `for` loop.
- Create `beta_hat_1` before writing the `for` loop. Make it a vector of length $1500$ where each element is `0`.
- Inside the body of the `for` loop, simulate new $y$ data each time. Use a variable to temporarily store this data together with the known $x$ data as a data frame.
- After simulating the data, use `lm()` to fit a regression. Use a variable to temporarily store this output.
- Use the `coef()` function and `[]` to extract the correct estimated coefficient.
- Use `beta_hat_1[i]` to store in elements of `beta_hat_1`.
- See the notes on [Distribution of a Sample Mean](http://daviddalpiaz.github.io/appliedstats/introduction-to-r.html#distribution-of-a-sample-mean) for some inspiration.

You can do this differently if you like. Use of these hints is not required.

```{r}
beta_hat_1 = c(rep(0,1500))

for (i in 1:1500) {
  
  above = rnorm(25, 0, sigma)
  x = runif(n = 25, 0, 10)
  y = beta_0 + beta_1 * x + above
  df_fit = lm(y ~ x)
  
  beta_hat_1[i] = coef(df_fit)[2]
}
```

**(e)** Report the mean and standard deviation of `beta_hat_1`. Do either of these look familiar?

Mean of $\hat{\beta_1}$ is: `r mean(beta_hat_1)`
Standard deviation of $\hat{\beta_1}$ is: `r sd(beta_hat_1)`

Yes, values look familiar especially the mean that is equal to the {\beta_1}

**(f)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.

```{r}
hist(beta_hat_1)
```
The shape is Symmetry meaning that the media and the mean are equal or very close 

***

## Exercise 4 (Be a Skeptic)

Consider the model

\[
Y_i = 3 + 0 \cdot x_i + \epsilon_i
\]

with

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 4)
\]

where $\beta_0 = 3$ and $\beta_1 = 0$.

Before answering the following parts, set a seed value equal to **your** birthday, as was done in the previous exercise.

```{r}
birthday = 19900826
set.seed(birthday)
```

**(a)** Use `R` to repeat the process of simulating `n = 75` observations from the above model $2500$ times. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
beta_0 = 3
beta_1 = 0
sigma = sqrt(4)
beta_hat_1 = c(rep(0, 2500))

for (i in 1:2500){
  above = rnorm(75, 0, sigma)
  x = runif(n = 75, 0, 10)
  y = beta_0 + beta_1 * x + above
  df_fit = lm(y ~ x)
  beta_hat_1[i] = coef(df_fit)[2]
}

```

Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Hint: Yes $\beta_1 = 0$.

**(b)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.
```{r}
hist(beta_hat_1)
```

The majority of the data are on the left side the Histogram skewed to the right because the mean is higher than the media 

**(c)** Import the data in [`skeptic.csv`](skeptic.csv) and fit a SLR model. The variable names in `skeptic.csv` follow the same convention as those returned by `sim_slr()`. Extract the fitted coefficient for $\beta_1$.

```{r}
library(readr)
skeptic = read_csv("skeptic.csv", 
    col_types = cols(predictor = col_double(), 
        response = col_double()))
```

```{r}
skeptic_df = lm(skeptic$response ~ skeptic$predictor, skeptic)
beta_1_hat_skeptic = coef(skeptic_df)[2]
beta_1_hat_skeptic
```

**(d)** Re-plot the histogram from **(b)**. Now add a vertical red line at the value of $\hat{\beta_1}$ in part **(c)**. To do so, you'll need to use `abline(v = c, col = "red")` where `c` is your value.

```{r}
hist(beta_hat_1)
abline(v = beta_1_hat_skeptic, col = "red")
```


**(e)** Your value of $\hat{\beta_1}$ in **(c)** should be negative. What proportion of the `beta_hat_1` values is smaller than your $\hat{\beta_1}$? Return this proportion, as well as this proportion multiplied by `2`.

```{r}
proportion = sum(beta_hat_1 < beta_1_hat_skeptic)
proportion
proportion * 2
```


**(f)** Based on your histogram and part **(e)**, do you think the [`skeptic.csv`](skeptic.csv) data could have been generated by the model given above? Briefly explain.

The $\hat{\beta_1}$ is almost the same as the model, this model covers a good areas of the chart following the empirical rule. 

***

## Exercise 5 (Comparing Models)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will perform some data cleaning before proceeding.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

We have:

- Loaded the data from the package
- Subset the data to relevant variables
    - This is not really necessary (or perhaps a good idea) but it makes the next step easier
- Given variables useful names
- Removed any observation with missing values
    - This should be given much more thought in practice

For this exercise we will define the "Root Mean Square Error" of a model as

\[
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}.
\]

**(a)** Fit three SLR models, each with "ozone" as the response. For the predictor, use "wind speed," "humidity percentage," and "temperature" respectively. For each, calculate $\text{RMSE}$ and $R^2$. Arrange the results in a markdown table, with a row for each model. Suggestion: Create a data frame that stores the results, then investigate the `kable()` function from the `knitr` package.

```{r}
library("knitr")
fit_wind = lm(Ozone$ozone ~ Ozone$wind, data = Ozone)
fit_humidity = lm(Ozone$ozone ~ Ozone$humidity, data = Ozone)
fit_temp = lm(Ozone$ozone ~ Ozone$temp, data = Ozone)

RMSE = function(error) { sqrt(mean(error^2)) }
wind_RMSE = RMSE(fit_wind$residuals)
humidity_RMSE = RMSE(fit_humidity$residuals)
temp_RMSE = RMSE(fit_temp$residuals)

R_wind =summary(fit_wind)$r.squared
R_humidity = summary(fit_humidity)$r.squared
R_tem = summary(fit_temp)$r.squared

df5 = data.frame("Model" = c("Wind:", "Humidity:", "Temp:"), "R^2" =  c(R_wind, R_humidity, R_tem), 
           "RMSE" = c(wind_RMSE, humidity_RMSE, temp_RMSE))
kable(df5, "markdown", align = "lrr", caption = "Model Table")
```


**(b)** Based on the results, which of the three predictors used is most helpful for predicting ozone readings? Briefly explain.

The Temp Model because it has better $R^2$ and lower $\text{RMSE}$

***

## Exercise 00 (SLR without Intercept)

**This exercise will _not_ be graded and is simply provided for your information. No credit will be given for the completion of this exercise. Give it a try now, and be sure to read the solutions later.**

Sometimes it can be reasonable to assume that $\beta_0$ should be 0. That is, the line should pass through the point $(0, 0)$. For example, if a car is traveling 0 miles per hour, its stopping distance should be 0! (Unlike what we saw in the book.)

We can simply define a model without an intercept,

\[
Y_i = \beta x_i + \epsilon_i.
\]

**(a)** [In the **Least Squares Approach** section of the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#least-squares-approach) you saw the calculus behind the derivation of the regression estimates, and then we performed the calculation for the `cars` dataset using `R`. Here you need to do, but not show, the derivation for the slope only model. You should then use that derivation of $\hat{\beta}$ to write a function that performs the calculation for the estimate you derived. 

In summary, use the method of least squares to derive an estimate for $\beta$ using data points $(x_i, y_i)$ for $i = 1, 2, \ldots n$. Simply put, find the value of $\beta$ to minimize the function

\[
f(\beta)=\sum_{i=1}^{n}(y_{i}-\beta x_{i})^{2}.
\]

Then, write a function `get_beta_no_int` that takes input:

- `x` - A predictor variable
- `y` - A response variable

The function should then output the $\hat{\beta}$ you derived for a given set of data.

**(b)** Write your derivation in your `.Rmd` file using TeX. Or write your derivation by hand, scan or photograph your work, and insert it into the `.Rmd` as an image. See the [RMarkdown documentation](http://rmarkdown.rstudio.com/) for working with images.

**(c)** Test your function on the `cats` data using body weight as `x` and heart weight as `y`. What is the estimate for $\beta$ for this data?

**(d)** Check your work in `R`. The following syntax can be used to fit a model without an intercept:

```{r, eval = FALSE}
lm(response ~ 0 + predictor, data = dataset)
```

Use this to fit a model to the `cat` data without an intercept. Output the coefficient of the fitted model. It should match your answer to **(c)**.

