---
title: "Homework 4"
author: "Omar Boffil"
date: "6/13/2020"
output: html_document
---

## Exercise 1 (Using `lm`)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` 
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

**(a)** Fit the following multiple linear regression model in `R`. Use `Calories` as the response and `Fat`, `Sugar`, and `Sodium` as predictors.

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \epsilon_i.
\]

Here,

- $Y_i$ is `Calories`.
- $x_{i1}$ is `Fat`.
- $x_{i2}$ is `Sugar`.
- $x_{i3}$ is `Sodium`.

```{r}
library(readr)
nutrition_2018 = read.csv("nutrition-2018.csv")
nmodel = lm(Calories ~ Fat + Sugar + Sodium, data = nutrition_2018)
null_value = lm(Calories ~ 1, data = nutrition_2018)
```


Use an $F$-test to test the significance of the regression. Report the following:
 
- The null and alternative hypotheses

  H0: β1=β2=β2...β(P - 1) = 0

  H1: At least one of βj ≠ 0, j = 1, 2, ⋯ , (P - 1)

- The value of the test statistic is `r anova(null_value, nmodel)[2,5]`

- The p-value of the test is: `r anova(null_value, nmodel)[2,6]`

- A statistical decision at $\alpha = 0.01$ is that we reject the hypothesis because the P values is less than alpha which is `r anova(null_value, nmodel)[2,6]` 

- A conclusion in the context of the problem in that exists a significant linear relationship between the response (Calorie) and at least one predictor selected 


When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**(b)** Output only the estimated regression coefficients. Interpret all $\hat{\beta}_j$ coefficients in the context of the problem.

```{r}
summary(nmodel)$coefficients[,1]
```
This means that for a value of Fat and sugar while sodium ingestion increases, Calories increases by 6.165246e-03, and the same explanation apply for the other coefficients, also the values are all positive, so the more you increase the amounts of Fat, sugar or sodium the more calories you will add to the food

**(c)** Use your model to predict the number of `Calories` in a Big Mac. According to [McDonald's publicized nutrition facts](https://www.mcdonalds.com/us/en-us/about-our-food/nutrition-calculator.html), the Big Mac contains 30g of fat, 9g of sugar, and 1010mg of sodium.

```{r}
Macd = data.frame(Fat = 30, Sugar = 9, Sodium = 1010)
predict(nmodel, newdata = Macd)


```
The number of Calories in a Big Mac is `r predict(nmodel, newdata = Macd)`

**(d)** Calculate the standard deviation, $s_y$, for the observed values in the Calories variable. Report the value of $s_e$ from your multiple regression model. Interpret both estimates in the context of this problem.


The $s_y$ is: `r sd(nutrition_2018$Calories)`

The $s_e$ is: `r summary(nmodel)$sigma`

The result tells us that if we consider Fat, Sugar and sodium the real deviation for calories is 80.8543 from the calories means  


**(e)** Report the value of $R^2$ for the model. Interpret its meaning in the context of the problem.

The $R^2$ is: `r summary(nmodel)$r.squared`, and this the proportion of the variance for Calories that is explained by Sugar, FAt, and Sodium 


**(f)** Calculate a 90% confidence interval for $\beta_2$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(nmodel, level = .90)["Sugar", ]
```
We are 90% confident that for the value of Fat and Sodium, the mean value of Sugar is between 3.783051 and 4.017983.

**(g)** Calculate a 95% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(nmodel, level = .95)["(Intercept)", ]
```
We are 95% confident that for 0 of Fat, sugar, and Sodium, the mean value of Calories is between 97.69443 and 103.21768 

**(h)** Use a 99% confidence interval to estimate the mean Calorie content of a food with 23g of fat, 0g of sugar, and 400mg of sodium, which is true of a large order of McDonald's french fries. Interpret the interval in context.
```{r}

true_order = data.frame(Fat = 23, Sugar = 0, Sodium = 400)
predict(nmodel, newdata = true_order, interval = "confidence", level = .99)

```
We are 99% confidence that for a large order of McDonald's, the means expected Calories is 298.0378, and also, we are confidence that the mean would will be between 294.3532 and 301.7224.
 
**(i)** Use a 99% prediction interval to predict the Calorie content of a Crunchwrap Supreme, which has 21g of fat, 6g of sugar, and 1200mg of sodium according to [Taco Bell's publicized nutrition information](https://www.tacobell.com/nutrition/info). Interpret the interval in context.
```{r}

Crunchwrap = data.frame(Fat = 21, Sugar = 6, Sodium = 1200)
predict(nmodel, newdata = Crunchwrap, interval = "prediction", level = .99)

```
We predict that Calories value would be 309.4065 and it would oscillate between 101.0345 and 517.7786

***

## Exercise 2 (More `lm` for Multiple Regression)

For this exercise we will use the data stored in [`goalies.csv`](goalies.csv). It contains career data for 462 players in the National Hockey League who played goaltender at some point up to and including the 2014-2015 season. The variables in the dataset are:
 
- `W` - Wins
- `GA` - Goals Against
- `SA` - Shots Against
- `SV` - Saves
- `SV_PCT` - Save Percentage
- `GAA` - Goals Against Average
- `SO` - Shutouts
- `MIN` - Minutes
- `PIM` - Penalties in Minutes

For this exercise we will consider three models, each with Wins as the response. The predictors for these models are:

- Model 1: Goals Against, Saves
- Model 2: Goals Against, Saves, Shots Against, Minutes, Shutouts
- Model 3: All Available

```{r}
goalies = read_csv("goalies.csv")
model1 = lm(W ~ GA + SV, data = goalies)
model2 = lm(W ~ GA + SV + SA + MIN + SO, data = goalies)
model3 = lm(W ~ ., data = goalies)
```

**(a)** Use an $F$-test to compares Models 1 and 2. Report the following:

```{r}
anova(model1, model2)
```


- The null hypothesis

**(H0:)** βsa = βmin = βso = 0

**(H1:)** At least one of βj ≠ 0

- The value of the test statistic is `r anova(model1, model2)[2, 5]`

- The p-value of the test is `r anova(model1, model2)[2,6]`

- A statistical decision at $\alpha = 0.05$ is that we reject the null Hypothesis

- The model you prefer
  
  I will prefer model 2

**(b)** Use an $F$-test to compare Model 3 to your preferred model from part **(a)**. Report the following:

```{r}

anova(model2, model3)

```

- The null hypothesis

**(H0:)** βsvpct = βgaa = βpim = 0

**(H1:)** At least one of βj ≠ 0

- The value of the test statistic is `r anova(model2, model3)[2, 5]`

- The p-value of the test is `r anova(model2, model3)[2,6]`

- A statistical decision at $\alpha = 0.05$ is that we reject the null Hypothesis

- The model you prefer

I will prefer the model 2



**(c)** Use a $t$-test to test $H_0: \beta_{\texttt{SV}} = 0 \ \text{vs} \ H_1: \beta_{\texttt{SV}} \neq 0$ for the model you preferred in part **(b)**. Report the following:

```{r}
summary(model2)
```

- The value of the test statistic is `r summary(model2)$coefficients[3, 3]`

- The p-value of the test `r summary(model2)$coefficients[3, 4]`

- A statistical decision at $\alpha = 0.05$ is that we reject the null Hypothesis

***

## Exercise 3 (Regression without `lm`)

For this exercise we will once again use the `Ozone` data from the `mlbench` package. The goal of this exercise is to fit a model with `ozone` as the response and the remaining variables as predictors.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
ozonefit = lm(ozone ~ ., data = Ozone)

```

**(a)** Obtain the estimated regression coefficients **without** the use of `lm()` or any other built-in functions for regression. That is, you should use only matrix operations. Store the results in a vector `beta_hat_no_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_no_lm ^ 2)`.

```{r}
n = nrow(Ozone)
X = cbind(rep(1, n), Ozone$wind, Ozone$humidity, Ozone$temp)
y = Ozone$ozone
beta_hat = solve(t(X) %*% X) %*% t(X) %*% y
beta_hat_no_lm = as.vector(beta_hat)
beta_hat_no_lm
sum(beta_hat_no_lm ^ 2)

```
**(b)** Obtain the estimated regression coefficients **with** the use of `lm()`. Store the results in a vector `beta_hat_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_lm ^ 2)`.

```{r}
ozonefit = lm(ozone ~ ., data = Ozone)
beta_hat_lm = as.vector(coef(ozonefit))
beta_hat_lm
sum(beta_hat_lm ^ 2)
```


**(c)** Use the `all.equal()` function to verify that the results are the same. You may need to remove the names of one of the vectors. The `as.vector()` function will do this as a side effect, or you can directly use `unname()`.

```{r}
all.equal(beta_hat_no_lm, beta_hat_lm)
```


**(d)** Calculate $s_e$ without the use of `lm()`. That is, continue with your results from **(a)** and perform additional matrix operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
p = ncol(Ozone)
y_hat = X %*% solve(t(X) %*% X) %*% t(X) %*% y
e = y - y_hat
Se = sqrt(t(e) %*% e / (n - p))
Se = as.vector(Se)
Selm = as.vector(summary(ozonefit)$sigma)
all.equal(Se, Selm)

Se
Selm

```



**(e)** Calculate $R^2$ without the use of `lm()`. That is, continue with your results from **(a)** and **(d)**, and perform additional operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
SSR = sum((y_hat - mean(y)) ^ 2)
SST = sum((y - mean(y))^2)
R_2 = as.vector(SSR/SST)
R_2lm = as.vector(summary(ozonefit)$r.squared)
all.equal(R_2, R_2lm)
R_2
R_2lm
```
***

## Exercise 4 (Regression for Prediction)

For this exercise use the `Auto` dataset from the `ISLR` package. Use `?Auto` to learn about the dataset. The goal of this exercise is to find a model that is useful for **predicting** the response `mpg`. We remove the `name` variable as it is not useful for this analysis. (Also, this is an easier to load version of data from the textbook.)

```{r}
# load required package, remove "name" variable
library(ISLR)
Auto = subset(Auto, select = -c(name))
```

When evaluating a model for prediction, we often look at RMSE. However, if we both fit the model with all the data as well as evaluate RMSE using all the data, we're essentially cheating. We'd like to use RMSE as a measure of how well the model will predict on *unseen* data. If you haven't already noticed, the way we had been using RMSE resulted in RMSE decreasing as models became larger.

To correct for this, we will only use a portion of the data to fit the model, and then we will use leftover data to evaluate the model. We will call these datasets **train** (for fitting) and **test** (for evaluating). The definition of RMSE will stay the same

\[
\text{RMSE}(\text{model, data}) = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}
\]

where

- $y_i$ are the actual values of the response for the given data.
- $\hat{y}_i$ are the predicted values using the fitted model and the predictors from the data.

However, we will now evaluate it on both the **train** set and the **test** set separately. So each model you fit will have a **train** RMSE and a **test** RMSE. When calculating **test** RMSE, the predicted values will be found by predicting the response using the **test** data with the model fit using the **train** data. *__Test__ data should never be used to fit a model.*

- Train RMSE: Model fit with *train* data. Evaluate on **train** data.
- Test RMSE: Model fit with *train* data. Evaluate on **test** data.

Set a seed of `11`, and then split the `Auto` data into two datasets, one called `auto_trn` and one called `auto_tst`. The `auto_trn` data frame should contain 292 randomly chosen observations. The `auto_tst` data will contain the remaining observations. Hint: consider the following code:

```{r}
set.seed(11)
auto_trn_idx = sample(1:nrow(Auto), 292)
auto_trn = Auto[auto_trn_idx,]
auto_tst = Auto[-auto_trn_idx,]
```

Fit a total of five models using the training data.

- One must use all possible predictors.
- One must use only `displacement` as a predictor.
- The remaining three you can pick to be anything you like. One of these should be the *best* of the five for predicting the response.

```{r}
library(knitr)
N = 292
N2 = 100

full_model = lm(mpg ~ ., data = auto_trn)
displacement_model = lm(mpg ~ displacement, data = auto_trn)
trn_model3 = lm(mpg ~ cylinders + displacement + horsepower + weight, data = auto_trn)
trn_model4 = lm(mpg ~ cylinders + horsepower + weight + year, data = auto_trn)
trn_model5 = lm(mpg ~ cylinders + horsepower + acceleration + year, data = auto_trn)


RMSE_full_model = sqrt(sum(residuals(full_model)^2) / N)
RMSE_displacement_model = sqrt(sum(residuals(displacement_model)^2) / N)
RMSE_train_model_1 = sqrt(sum(residuals(trn_model3 )^2) / N)
RMSE_train_model_2 = sqrt(sum(residuals(trn_model4)^2) / N)
RMSE_train_model_3 = sqrt(sum(residuals(trn_model5)^2) / N)


RMSE_TEST_FULL = sqrt((sum(auto_tst$mpg - predict(full_model, newdata = auto_tst)) ^ 2)/ N2)
RMSE_displacement_TEST = sqrt((sum(auto_tst$mpg - predict(displacement_model, newdata = auto_tst)) ^ 2)/ N2)
RMSE_model3_TEST = sqrt((sum(auto_tst$mpg - predict(trn_model3, newdata = auto_tst)) ^ 2)/ N2)
RMSE_model4_TEST = sqrt((sum(auto_tst$mpg - predict(trn_model4, newdata = auto_tst)) ^ 2)/ N2)
RMSE_model4_TEST = sqrt((sum(auto_tst$mpg - predict(trn_model5, newdata = auto_tst)) ^ 2)/ N2)


table = data.frame(
  "Model" = c("Full_Model","Displacement_Model","Mode3","Model4","Model5"),
  "Train" = c(RMSE_full_model, RMSE_displacement_model, RMSE_train_model_1, RMSE_train_model_2, RMSE_train_model_3),
  "Test" = c(RMSE_TEST_FULL, RMSE_displacement_TEST, RMSE_model3_TEST, RMSE_model4_TEST, RMSE_model4_TEST))

kable(table, caption = "RMSE")
```

For each model report the **train** and **test** RMSE. Arrange your results in a well-formatted markdown table. Argue that one of your models is the best for predicting the response.

The Full model is the best to predict the response this because of the Trained models the one with the lowest RMSE is it with a value of `r RMSE_full_model` 

***

## Exercise 5 (Simulating Multiple Regression)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \beta_4 x_{i4} + \beta_5 x_{i5} + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 2$
- $\beta_1 = -0.75$
- $\beta_2 = 1.5$
- $\beta_3 = 0$
- $\beta_4 = 0$
- $\beta_5 = 2$
- $\sigma^2 = 25$

We will use samples of size `n = 42`.

We will verify the distribution of $\hat{\beta}_2$ as well as investigate some hypothesis tests.

**(a)** We will first generate the $X$ matrix and data frame that will be used throughout the exercise. Create the following nine variables:

- `x0`: a vector of length `n` that contains all `1`
- `x1`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `x2`: a vector of length `n` that is randomly drawn from a uniform distribution between `0` and `4`
- `x3`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `1`
- `x4`: a vector of length `n` that is randomly drawn from a uniform distribution between `-2` and `2`
- `x5`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `X`: a matrix that contains `x0`, `x1`, `x2`, `x3`, `x4`, and `x5` as its columns
- `C`: the $C$ matrix that is defined as $(X^\top X)^{-1}$
- `y`: a vector of length `n` that contains all `0`
- `sim_data`: a data frame that stores `y` and the **five** *predictor* variables. `y` is currently a placeholder that we will update during the simulation.

Report the sum of the diagonal of `C` as well as the 5th row of `sim_data`. For this exercise we will use the seed `420`. Generate the above variables in the order listed after running the code below to set a seed.

```{r}
set.seed(420)
sample_size = 42

beta_0 = 2
beta_1 = -.75
beta_2 = 1.5
beta_3 = 0
beta_4 = 0
beta_5 = 2
sigma = sqrt(25)

x0 = rep(1,sample_size)
x1 = rnorm(sample_size,0,2)
x2 = runif(sample_size,0,4)
x3 = rnorm(sample_size,0,1)
x4 = runif(sample_size,-2,2)
x5 = rnorm(sample_size, 0, 2)
X = cbind(x0, x1, x2, x3, x4, x5)
C = solve(t(X) %*% X)
y = rep(0,sample_size)
sim_data = data.frame(y, x1, x2, x3, x4, x5)

sum(diag(C))


```
```{r}
sim_data[5, ]
```

**(b)** Create three vectors of length `2500` that will store results from the simulation in part **(c)**. Call them `beta_hat_1`, `beta_3_pval`, and `beta_5_pval`.


```{r}
beta_hat_1 = c(rep(0, 2500))
beta_3_pval = c(rep(0, 2500))
beta_5_pval = c(rep(0, 2500))
```


**(c)** Simulate 2500 samples of size `n = 42` from the model above. Each time update the `y` value of `sim_data`. Then use `lm()` to fit a multiple regression model. Each time store:

- The value of $\hat{\beta}_1$ in `beta_hat_1`
- The p-value for the two-sided test of $\beta_3 = 0$ in `beta_3_pval`
- The p-value for the two-sided test of $\beta_5 = 0$ in `beta_5_pval`

```{r}
for(i in 1:2500) {
  e = rnorm(sample_size, mean = 0, sd = sigma)
  sim_data$y = beta_0 * x0 + beta_1 * x1 + beta_2 * x2 + beta_3 * x3 + beta_4 * x4 + beta_5 * x5 + e
  model_fit = lm(y ~ x1 + x2 + x3 + x4 + x5, data = sim_data)
  beta_hat_1[i] = coef(model_fit)[2]
  beta_3_pval[i] = summary(model_fit)$coefficients[4,4]
  beta_5_pval[i] = summary(model_fit)$coefficients[6,4]
}

```

**(d)** Based on the known values of $X$, what is the true distribution of $\hat{\beta}_1$?

```{r}
hist(beta_hat_1, breaks = 20)
```
It is a normal distribution where the majority of the values are negative and a mean of `r mean(beta_hat_1)`. 

**(e)** Calculate the mean and variance of `beta_hat_1`. Are they close to what we would expect? Plot a histogram of `beta_hat_1`. Add a curve for the true distribution of $\hat{\beta}_1$. Does the curve seem to match the histogram?
```{r}
mean(beta_hat_1)
var(beta_hat_1)
```
Yes, they are close that what we expect

```{r}

hist( beta_hat_1, prob = TRUE, breaks =30, main = "beta_hat_1")
curve(dnorm(x, mean = beta_1, sd = sqrt(sigma ^ 2 * C[1+1,1+1])), col = "orange", add = TRUE, lwd = 3)

```
The curve matches the histogram very well

**(f)** What proportion of the p-values stored in `beta_3_pval` is less than 0.10? Is this what you would expect?

```{r}
sum(beta_3_pval < 0.10) / 2500
```

This is not what we expected 

**(g)** What proportion of the p-values stored in `beta_5_pval` is less than 0.01? Is this what you would expect?
```{r}
sum(beta_5_pval < 0.01) / 2500
```
This is what we expected 
