---
title: "Week 3 Homework"
author: "Omar Boffil"
date: "6/6/2020"
output: html_document
---
## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

```{r}
library(MASS)
cat_model = lm(Hwt ~ Bwt, data = cats)
summary(cat_model)
```
- The null and alternative hypotheses
  H0:β1 = 0 vs H1:β1 ≠ 0
  
- The value of the test statistic is `r summary(cat_model)$coefficients[2,3]` for β1

- The p-value of the test is `r summary(cat_model)$coefficients[2,4]` for β1

- A statistical decision at $\alpha = 0.05$ is Reject null Hypotecys because the p-value value is very low (`r summary(cat_model)$coefficients[2,4]`)

- In conclusion, we can see a linear relationship between heart weight and body weight in the cast dataset according to the p-value obtained.

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**(b)** Calculate a 95% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, level = 0.95)[2,]
```

On average, we are 95% confident that cat heart weight increase between 3.539343 and 4.528782 for each Kg of cat body weight.

**(c)** Calculate a 90% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, level = 0.90)[1,]
```
On average, we are 90% confident that the weight of cat heart is between -1.5028345  and 0.7895096 for a cat of body weight 0 Kg, but we know that it is impossible to have a negative car heart weight.

**(d)** Use a 90% confidence interval to estimate the mean heart weight for body weights of 2.1 and 2.8 kilograms. Which of the two intervals is wider? Why?

```{r}
predict(cat_model, newdata = data.frame(Bwt = c(2.1, 2.8)), interval ="confidence", level = .90)
```

The 2.8 interval is wider because the length between the mean and the upper and lower bounds is larger than the ones for the range 2.1

**(e)** Use a 90% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.

```{r}
predict(cat_model, newdata = data.frame(Bwt = c(2.8, 4.2)), interval ="predict", level = .90)
```

**(f)** Create a scatterplot of the data. Add the regression line, 95% confidence bands, and 95% prediction bands.

```{r}
sequence_num = seq(min(cats$Bwt), max(cats$Bwt), by = 0.01)
cat_conf_line = predict(cat_model, newdata = data.frame(Bwt = sequence_num), interval = "confidence", level = 0.95)
cat_pred_line = predict(cat_model, newdata = data.frame(Bwt = sequence_num), interval = "prediction", level = 0.95)

plot(Hwt ~ Bwt, data = cats,
     xlab = "Body Weight (Kg)",
     ylab = "Heart Weigth (grams)",
     main = "Body weight vs Heart Weight",
     pch = 20,
     cex  = 1,
     col  = "black")
abline(cat_model, lwd = 1, col = "red")
lines(sequence_num, cat_conf_line[,2], col = "blue", lwd = 3, lty = 2)
lines(sequence_num, cat_conf_line[,3], col = "blue", lwd = 3, lty = 2)
lines(sequence_num, cat_pred_line[,2], col = "grey", lwd = 3, lty = 3)
lines(sequence_num, cat_pred_line[,3], col = "grey", lwd = 3, lty = 3)
```


**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

```{r}
beta1_hat = summary(cat_model)$coefficients[2,1]
SE = summary(cat_model)$coefficients[2,2]
t_value = (beta1_hat - 4) / SE
num_rows = nrow(cats)
p_value = pt(t_value, df = num_rows - 2, lower.tail = FALSE) * 2
t_value
p_value
```


- The value of the test statistic is: `r t_value`
- The p-value of the test is: `r p_value`
- A statistical decision at $\alpha = 0.05$ is fail to reject hypothesis because the p-values is higher than $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

```{r}
ozone_wind_model = lm(Ozone$ozone ~ Ozone$wind, data = Ozone)
summary(ozone_wind_model)
```


- The null and alternative hypotheses
  H0:β1 = 0 vs H1:β1 ≠ 0 
  
- The value of the test statistic is `r summary(ozone_wind_model)$coefficients[2,3]` for β1

- The p-value of the test is `r summary(ozone_wind_model)$coefficients[2,4]` for β1

- A statistical decision at $\alpha = 0.01$is that we fail to reject the null hypothesis because p-value is higher than alpha value

- In conclusion, base on our hypothesis, we didn't find any relationship between ozone measurement and wind speed. So, it is no linear relationship between these variables

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]


Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

```{r}
ozone_temp_model = lm(Ozone$ozone ~ Ozone$temp, data = Ozone)
summary(ozone_temp_model)
```
- The null and alternative hypotheses
  H0:β1 = 0 vs H1:β1 ≠ 0
  
- The value of the test statistic is `r summary(ozone_temp_model)$coefficients[2,3]` for β1

- The p-value of the test is `r summary(ozone_temp_model)$coefficients[2,4]` for β1

- A statistical decision at $\alpha = 0.01$ is Reject null Hypotecys because the p-value value is very low (`r summary(ozone_temp_model)$coefficients[2,4]`)

- In conclusion, we can see a linear relationship between ozone measurement and temperature according to the p-value obtained.

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19900826
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```
```{r}
beta_0 = -5
beta_1 = 3.25
sigma = sqrt(16)

beta_hat_0 = rep(0,2000)
beta_hat_1 = rep(0,2000)

for(i in 1:2000) {
  e = rnorm(n,mean=0,sd=sigma)
  y = beta_0 + beta_1 * x + e
  model = lm(y ~ x)
  beta_hat_0[i] = coef(model)[1]
  beta_hat_1[i] = coef(model)[2]
}

```


**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values

```{r}
library("knitr")
Sxx = sum((x-mean(x)) ^2)
var_beta_hat_1 = sigma ^ 2 / Sxx 
var_beta_hat_0 = sigma ^ 2 * ( 1/ n + mean(x) ^ 2 / Sxx)
Beta_0_values = c(beta_0, mean(beta_hat_0), sqrt(var_beta_hat_0), sd(beta_hat_0))
Beta_1_values = c(beta_1, mean(beta_hat_1), sqrt(var_beta_hat_1), sd(beta_hat_1))
const = c("True Mean", "Simulated mean", "True Stand. Devi.", "Simulated Stand. Devi")

df = data.frame("Values" = const, "Hat_β0" = Beta_0_values, "Hat_β1" = Beta_1_values)
kable(df, "markdown", align = "lrr", caption = "Expected Table")
```


**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.

```{r}
posi_tion = c(1, 2)
par(mfrow = posi_tion)
hist( beta_hat_1, prob = TRUE, breaks = 20, 
      xlab = expression(hat(beta)[1]), 
      ylab= "Values", 
      main = "Hat_β1" , border = "blue")
curve(dnorm(x, mean = beta_1, sd = sqrt(var_beta_hat_1)), col = "red", add = TRUE, lwd = 3) 

hist( beta_hat_0, prob = TRUE, breaks =20, 
      xlab = expression(hat(beta)[0]), 
      ylab = "Values",  
      main = "Hat_β0", border = "blue")
curve(dnorm(x, mean = beta_0, sd = sqrt(var_beta_hat_0)), col = "red", add = TRUE, lwd = 3)
```

***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19900826
set.seed(birthday)
n = 25
x = seq(0, 2.5, length = n)

beta_0 = 5
beta_1 = 2
sigma = sqrt(9)

beta_hat_0 = rep(0,2500)
beta_hat_1 = rep(0,2500)
Se = rep(0,2500)

for(i in 1:2500) {
  e = rnorm(n,mean=0,sd=sigma)
  y = beta_0 + beta_1 * x + e
  model = lm(y ~ x)
  beta_hat_0[i] = coef(model)[1]
  beta_hat_1[i] = coef(model)[2]
  Se[i] = summary(model)$coefficient[2,2]
}

```

**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.

```{r}
crit_95 = abs(qt(.05 / 2, df = n - 2))
lower_95 = c(beta_hat_1 - crit_95 * Se)
upper_95 = c(beta_hat_1 + crit_95 * Se)

values95 = data.frame("Hat_β1" = beta_hat_1, Lower = lower_95, Upper = upper_95)
head(values95)
```


**(c)** What proportion of these intervals contains the true value of $\beta_1$?

```{r}
mean(lower_95 < beta_1 & beta_1 < upper_95)
```


**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?

```{r}
1 - mean(lower_95 < 0 & 0 < upper_95)
```


**(e)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 99% confidence interval. Store the lower limits in a vector `lower_99` and the upper limits in a vector `upper_99`.

```{r}
crit_99 = abs(qt(.001 / 2, df = n - 2))
lower_99 = c(beta_hat_1 - crit_99 * Se)
upper_99 = c(beta_hat_1 + crit_99 * Se)
values99 = data.frame("Hat_β1" = beta_hat_1, Lower = lower_99, Upper = upper_99)
head(values99)
```


**(f)** What proportion of these intervals contains the true value of $\beta_1$?

```{r}
mean(lower_99 < beta_1 & beta_1 < upper_99)
```


**(g)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.01$?

```{r}
1 - mean(lower_99 < 0 & 0 < upper_99)
```

***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval

```{r}
calc_pred_int = function(model, newdata, level = 0.95)
{
  model = model
  newd_1 = newdata
  n = nrow(model$model)
  x1 = newd_1[1]
  lv = level
  
  x = model$model[names(newd_1)]
  
  xmean = mean(x[[1]])
  
  predic_va = predict(cat_model, newdata = x1)[1]
  
  Se = summary(model)$sigma
  
  beta_hat_0 = model$coefficients[1]
  
  beta_hat_1 = model$coefficients[2]
  
  alpha = 1 - lv
  
  crit_any_level = abs(qt(alpha / 2, df = nrow(model$model) - 2))
  
  Sxx = sum((x - mean(x[[1]])) ^ 2)
  
  lowert = predic_va - crit_any_level * Se * sqrt(1 + 1 / n + (x1 - xmean) ^ 2 / Sxx)
  uppert = predic_va + crit_any_level * Se * sqrt(1 + 1 / n + (x1 - xmean) ^ 2 / Sxx)
  
  return(c(lowert, uppert))
  
}

```


**(b)** After writing the function, run this code:

```{r}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```

**(c)** After writing the function, run this code:

```{r}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.90)
```


